---
title: "Eidith R Package Setup"
output: 
  rmarkdown::html_vignette
author: "Noam Ross"
vignette: >
  %\VignetteIndexEntry{Eidith R Package Setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  tidy = FALSE,
  error = FALSE,
  fig.width = 8,
  fig.height = 8)
```

The **eidith** R package provides programmatic access and analytical tools for
data from the [PREDICT Program](http://www.vetmed.ucdavis.edu/ohi/predict/),
housed at the [Emerging Infectious Disease Information Technology Hub](https://www.eidith.org/) (EIDITH).

The **eidith** package contains no data. To access data, you must be a
[registered](https://www.eidith.org/register.aspx) EIDITH user with data access
privileges. If you have a question about your access level, contact technology@eidith.org.

This package is currently in development. We anticipate additional functionality
for common analytical tasks. It also currently only accesses data from PREDICT-1. If
you have questions, bug reports, or feature requests, please post them on our
[issue tracker](https://github.com/ecohealthalliance/eidith/issues).

### Installation

**eidith** is installed from our own package repository rather than CRAN.  To
install the latest stable version, run the following command:

```{r install, eval = FALSE}
source("https://install-github.me/ecohealthalliance/eidith")
```

If you wish to install the latest *development* (unstable) version, you will
require the **devtools** package, like so:

```{r install-dev, eval = FALSE}
devtools::install_github('ecohealthalliance/eidith@dev')
```

### Authentication

To download data from EIDITH, you must provide logon credentials.  If you run
`ed_db_download()`, `ed_get()`, or `ed2_get()` functions from the R console, your EIDITH username
and password will be requested.  To use these functions in scripts, you 
must provide these credentials as *environment variables*.  

You can cache your credentials and login automatically by putting them in a hidden
`.Renviron` file, which defines environment variables for your R session. For
more information see the `ed_auth()` help file.

### Downloading data

The **eidith** package downloads data from EIDITH and stores it locally on your
computer in a database so you don't have to download from the web repeatedly.
Once you've installed the package, you'll need to populated this database with
the `ed_db_download()` command, like so:

```
library(eidith)
ed_db_download()
```
This may take a few minutes to complete depending on the speed of your internet
connection.

### Loading data

Once you've populated your local database, you can look up EIDITH data with
the *table functions*, each of which loads data from the different table in the
PREDICT-1 or PREDICT-2 database: `ed_events()`, `ed_animals()`, `ed_specimens()`, 
`ed2_events()`, `ed2_animals()` and so on. See the help for these functions for more
details.

`ed_metadata()` and `ed2_metadata()` contain lists of all variables across these tables
and some information on them.  If you lookup this in R help in RStudio with
`?ed_metadata` or `?ed2_metadata` you'll find this metadata in searchable form.
